---
title: "Working with Remote Zarr archives"
author: 
- name: Marcel Ramos
  affiliation: Roswell Park Comprehensive Cancer Center, Buffalo, NY
date: "`r Sys.Date()`"
output:
    BiocStyle::html_document:
        toc: true
        toc_float: true
package: ZarrExperiment
vignette: >
  %\VignetteIndexEntry{Working with Remote Zarr archives}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEndcoding{UTF-8}
---


# Using OSN on TMA Data

The Open Storage Network has allowed the Bioconductor project expanded
offerings to cloud data. Here we demonstrate the usage of the `RemoteZarr`
class to access Zarr data.

```{r}
library(ZarrExperiment)
```

# Data Location

https://mghp.osn.xsede.org/bir190004-bucket01/index.html#TMA11/zarr/

# Data Representation

## ZarrRemote class

```{r}
showClass("ZarrRemote")
```

## endpoint and bucket inputs

The `endpoint` corresponds to the OSN website and the `bucket` input corresponds
to the bucket name and subfolders where the `.zarr` archives can be located.

```{r}
zr <- ZarrRemote(
    endpoint = "https://mghp.osn.xsede.org/",
    bucket = "bir190004-bucket01/TMA11/zarr/"
)
zr
```

## URL input

A URL may also be used as the `resource` input. Both the `endpoint` and the
`bucket` location will be deduced when possible.

```{r}
zr <- ZarrRemote(
    resource = "https://mghp.osn.xsede.org/bir190004-bucket01/TMA11/zarr/"
)
zr
```

# Data Retrieval

Currently, data retrieval is possible on a single directory via the `import`
method on the `ZarrRemote` class.

## ZarrRemote `import` method

The `import` method makes internal use of the `s3fs` python module:

```{r}
getMethod(import, "ZarrRemote")
```

This operation takes a few minutes to complete and produces approximately a 2GB
array.

```{r,eval=FALSE}
import(zr, filename = "99.zarr")
```

# Session Info

```{r}
sessionInfo()
```


