---
title: "Working with Zarr archives"
author: 
- name: Martin Morgan
  affiliation: Roswell Park Comprehensive Cancer Center, Buffalo, NY
date: "`r Sys.Date()`"
output:
    BiocStyle::html_document:
        toc: true
        toc_float: true
package: ZarrExperiment
vignette: >
  %\VignetteIndexEntry{Working with Zarr archives}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEndcoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    cache = TRUE
)
```

# Introduction

The ZarrExperiment package is currently under construction.  The
purpose of the package is to build an interface between
zarr(https://zarr.readthedocs.io/en/stable/) and the Bioconductor
`SummarizedExperiment` family.

# Installation

Install the package with

```
R -e "BiocManager::install('Bioconductor/ZarrExperiment')"
```

The `ZarrExperiment` package uses the zarr module from
python. Configuring python currently requires a git clone.

```
git clone https://github.com/Bioconductor/ZarrExperiment
```

We suggest using a python virtual environment to configure
python. There are different ways of establishing virtual environments.

## `venv`

## `virtualenv`

`virtualenv` is a distinct program. Do these from the command line.

1) Install `python3` and `virtualenv`

2) Create the virtual environment.

   ```
   virtualenv -p python3 ~/.virtualenvs/Bioconductor
   ``` 

3) Activate this virtual environment.

   ```
   source ~/.virtualenvs/Bioconductor/bin/activate
   ```

4) Install the required python packages from the file
   `python-requirements.txt` in the package's base directory using
   pip.

   ```
   pip install -r python-requirements.txt
   ```

   When done, deactivate the virtual environment.

   ```
   deactivate
   ```

6) Set the enviroment `RETICULATE_PYTHON` variable to use the `Bioconductor`
   virtual environment.

   ```
   export RETICULATE_PYTHON=~/.virtualenvs/Bioconductor/bin/python
   ```

7) Test that `zarr` can be imported with reticulate.

   ```
   R -e "reticulate::import('zarr')"
   ```

## basilisk

# Use

Load libraries used in this vignette

```{r, message=FALSE}
library(SummarizedExperiment)
library(tibble)
library(ZarrExperiment)
```

Point to a sample archive. Archives are folders with a collection of files.

```{r}
fl <- system.file(
    package="ZarrExperiment", "extdata",
    "stahl-2016-science-olfactory-bulb.matrix.zarr"
)
dir(fl, recursive = TRUE, all.files = TRUE)
```

Load the archive and view, via the `show()` or `tree()` method, the
available groups (datasets) in the archive. These groups are analogous
to hdf5 datasets.

```{r}
arr <- ZarrArchive(fl)
arr
```

The archive allows `$` subsetting, including tab completion. Access
the `matrix` group member and coerce it to an _R_ (dense) matrix.

```{r}
m <- t(as(arr$matrix, "matrix"))
m[1:5, 1:5]
```

The components of this archive (archive format is completely general,
so it is not possible to write a 'smart' import function) can be
accessed...

```{r}
rowData <- tibble(
    gene_name = as(arr$gene_name, "matrix")
)
rowData

colData <- tibble(
    region_id = as(arr$region_id, "matrix"),
    x_region = as(arr$x_region, "matrix"),
    y_region = as(arr$y_region, "matrix")
)
colData
```

Form a `SummarizedExperiment` from these components:

```{r}
se <- SummarizedExperiment(
    assays = list(count = m),
    rowData = rowData,
    colData = colData
)
se
```

The `SummarizedExperiment` object can then be used in standard _R_ /
_Bioconductor_ single-cell and other work flows.

# Acknowledgements

```{r}
sessionInfo()
```
